---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

## SPCtools

SPCtools is an R package for clear, reproducible statistical process control (SPC) calculations. It provides opinionated, human‑readable functions for XmR, XbarR, ANOX, and ANOM analyses, designed for real operational use in regulated manufacturing and industrial statistics. It directly addresses gaps I have seen in existing commercial SPC software and workflows, such as rich metadata.

The package emphasizes:

- **Correct calculations** - SPCtools includes tests and edge‑case handling to ensure accurate limits and diagnostics across a wide range of data scenarios.
- **Reproducible workflows** - Functions return structured objects containing metadata about how limits were calculated, what baseline data was used, and why. This supports documentation, auditability, and long‑term monitoring.
- **Clarity over complexity** - SPC calculations are not computationally heavy, so SPCtools prioritizes readable code, minimal dependencies, and transparent logic.
- **Practical extensions** - In addition to standard control charts, SPCtools will includes ANOX and ANOM methods for fixed datasets, enabling intuitive diagnostics and exploratory analysis beyond ongoing monitoring.

## Why SPCtools?

I’ve used SPC methods for many years in manufacturing, R&D, and industrial analytics. Control charts are powerful tools for understanding variation, diagnosing problems, and guiding decisions. They are also essential prerequisites for many analyses - especially when data may show patterns over time, across batches, or during experiments.

Existing R packages occasionally produce inconsistent results or lack the workflow structure needed for documentation and reproducibility. SPCtools aims to provide:

- disciplined, correct calculations  
- metadata‑rich outputs  
- opinionated defaults that reflect real‑world usage  
- extensions for ANOX and ANOM  
- a foundation for future SPC‑related utilities  

## Installation

You can install the development version of SPCtools from GitHub:

```{r}
# install.packages("devtools")
# devtools::install_github("benjepson/SPCtools")
```

## Examples

Examples and vignettes are in development as the core functions and tests are completed. A pkgdown documentation site will be available soon.

## Roadmap

 - XmR and XbarR calculations with metadata
 - XmR is what I find most useful, so it is the first priority
 - Rare‑events XmR support
 - ANOX and ANOM functions (analogues of XmR and XbarR for one time analysis vs. ongoing monitoring)
 - Workflow helpers for documentation and reproducibility
 - Comprehensive test suite for edge‑case scenarios
 - pkgdown site with examples and guidance


